---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# validatEHR

validatEHR is a working version of the R package for the paper, "Integrating EHR Subject to Outcome Misclassification with Validation Data Subject to Selection Bias." The package implements two approaches for estimating the average treatment effect (ATE) when validation data containing gold standard outcomes are available.

## Installation

You can install the development version of validatEHR from ...

``` r
# install.packages("devtools")
devtools::install_github("jshen650/validatEHR")
```

## Example
This is a basic example which shows how to generate a data set from one of the simulation scenarios described in the paper and apply one of the two approaches described in the paper that leverages information from a validation data set containing gold standard outcomes.

```{r example, include=FALSE}
library(validatEHR)
```

Data can be simulated with the `genDat` function. This function requires specification of the following:  
* n = data set size  
* p11 = sensitivity of the outcomes  
* p10 = (1-specificity) of the outcomes  
* valType = type of validation data (e.g. "SRS" or "non-probability")  
* alphaInt = parameter for validation data set size; if not specified, the default will generate size of ~850  
* seedNum = seed number

```{r example data, include=TRUE}
## Simulate data following paper specifications - say n=5000, with non-probability validation sample of size nV~850

example_dat <- validatEHR::genDat(n=5000, p11=0.67, p10=0.24, valType="non-probability", seedNum=215)
str(example_dat)
```

The estimate of the ATE and its variance using gold standard outcomes from the validation data and silver standard outcomes from the non-validation individuals can be obtained using `run_method2`, which follows from applying Equation 2 in the paper. Alternatively, for using all silver standard outcomes rather than just a subset, estimation of the ATE and its variance can come from implementing `run_method4`. The user also has the option to use the optimal choice of weight, $w$, which leads to minimum variance for obtaining the final estimate of the ATE.

```{r example method, include=TRUE}
## estimate ATE and variance when using gold standard outcomes from validation sample
## and all silver standard outcomes

## specify relevant variables for treatment propensity model
varTrtMod = c("X1", "X2", "X3", "X4", "X5")

## specify relevant variables for validation sample selection model - 
## here, includes the same covariates as varTrtMod
varSelectMod = varTrtMod

example_ATE <- validatEHR::run_est4(example_dat, inVal=example_dat$inVal, varTrt="trt",
                                       varGold="y", varSilver="yStar", varTrtMod=varTrtMod, 
                                       varSelectMod = varSelectMod, opt="TRUE")

## display optimal choice of weight w, final estimate of ATE, final variance estimate of ATE,
# estimate of ATE using validation data, variance estimate of ATE using validation data,
# estimate of ATE using silver standard outcomes,
# and variance estimate of ATE using silver standard outcomes
example_ATE

```

